---
title: "Activity Monitoring"
author: "Biruk H. Sahle"
date: "December 2, 2016"
output: html_document
keep_md: true
---
##Loading and preprocessing the data
First, Download the "Activity.csv" file form course repository and set the working directory as the floder containing the "activity.csv" file. 

``` {r, echo = TRUE}
Activity <- read.csv("Activity.csv")
Activity$steps <- as.numeric(Activity$steps)
```

##What is mean total number of steps taken per day?

```{r, echo = TRUE}
Totalsteps<- aggregate(steps~date, Activity, sum, rm.na = TRUE)

hist(Totalsteps$steps, main = "Histogram of the total number of steps taken each day", xlab = "Total number of steps per day", col = "blue")

summary(Totalsteps$steps)    
```

The mean and median of the total number of steps taken per day is **10770** steps. 

## What is the average daily activity pattern?

```{r, echo= TRUE}
Totalstepsinterval <- read.csv("Activity.csv")
Totalstepsinterval$steps <- as.numeric(Totalstepsinterval$steps)
Totalstepsinterval <- aggregate(steps~interval, Totalstepsinterval, mean, na.rm = TRUE)
plot(Totalstepsinterval$interval, Totalstepsinterval$steps, type = "l", xlab = "Interval", ylab = "Average steps")

Totalstepsinterval$interval[which.max(Totalstepsinterval$steps)]

```

The **835th** interval is the 5 minute interval which contains the maximum number of steps on average. 

## Imputing missing values


```{r, echo= TRUE}
Activitymeancor <- read.csv("Activity.csv")
Activitymeancor$steps <- as.numeric(Activitymeancor$steps)
sum(is.na(Activitymeancor$steps))
paste(100*sum(is.na(Activitymeancor$steps))/length(Activitymeancor$steps),"%",sep="")
```

In total there are about **2304** missing values in the data set which is about 13.12% of the total data. 

The method used to impute the missings values is predictive mean matching. 

``` {r, echo = TRUE}
library(mice)
tempData1 <- mice(Activitymeancor,m=1,maxit=50,meth='pmm',seed=500)
Activitymeancor <- complete(tempData1,1)

Totalstepscor<- aggregate(steps~date, Activitymeancor, FUN = sum, rm.na = TRUE )

hist(Totalstepscor$steps, main = "Histogram of the total number of steps taken each day", xlab = "Total number of steps per day", col = "blue", freq = TRUE)

summary(Totalstepscor$steps) 

```

After imputing missing values the median (**10600**)and the mean (**10730**) of the total steps taken each day has decreased slightly. 

## Are there differences in activity patterns between weekdays and weekends?

``` {r, Echo= TRUE}
library(ggplot2)
Activitymeancorwd <- Activitymeancor
Activitymeancorwd$date <- as.Date(Activitymeancorwd$date,"%Y-%m-%d")
Activitymeancorwd$dofw <- weekdays(Activitymeancorwd$date)
Activitymeancorwd$dofw <- gsub("Monday", "Weekday", Activitymeancorwd$dofw, ignore.case = FALSE)
Activitymeancorwd$dofw <- gsub("Tuesday", "Weekday", Activitymeancorwd$dofw, ignore.case = FALSE)
Activitymeancorwd$dofw <- gsub("Wednesday", "Weekday", Activitymeancorwd$dofw, ignore.case = FALSE)
Activitymeancorwd$dofw <- gsub("Thursday", "Weekday", Activitymeancorwd$dofw, ignore.case = FALSE)
Activitymeancorwd$dofw <- gsub("Friday", "Weekday", Activitymeancorwd$dofw, ignore.case = FALSE)
Activitymeancorwd$dofw <- gsub("Saturday", "Weekend", Activitymeancorwd$dofw, ignore.case = FALSE)
Activitymeancorwd$dofw <- gsub("Sunday", "Weekend", Activitymeancorwd$dofw, ignore.case = FALSE)


Totalstepscorwd <- aggregate(steps~interval+dofw, Activitymeancorwd, mean)

qplot(interval, steps, data = Totalstepscorwd, facets = dofw ~., geom = "line", xlab = "Interval", ylab = "Average steps")
```



